\ZpL: a p-adic precision package

### 2018-02-03 ZpL/ZpL.tex
Add a conclusion

If provably correct results are desired, interval arithmetic provides an alternative to floating point.  Instead of just tracking an approximation to the answer, the package also tracks a radius within which the true result lies.  This method is commonly used for $p$-adic computations since the ultrametric property of $p$-adic fields frequently keeps the radius small.  Computations remain fairly efficient with this approach, but numerical instability can still lead to dramatic losses in precision, (see Section \ref{sec:demo} for many examples).

@@

Unfortunately, the latter condition cannot be checked on the matrix 
$M_t$ (because of reductions). 
However it is possible (and easy) to check whether the weaker condition 
$d\Phi_{t,\ttv}(K_t) \subsetneq p^N \Zp$. This checking is achieved by 
the method
{\color{method}\verb?is_precision_capped?} (provided by our package) which returns true 
if $d\Phi_{t,\ttv}(K_t) = p^N \Zp$. As a consequence, when this method
answers false, the absolute precision computed by the software is
likely optimal.
\end{rmk}

@@

In the \ZpLC framework, creating a variable has a cost $O(m)$. The case 
of deletion is more difficult to handle. However, by the temporal 
locality principle, it seems safe to assume that it is not the 
bottleneck (which is the case in practice). Therefore, when executed 
with the \ZpLF mecanism, the cost of our algorithm is expected to be 
roughly $O(mc)$. Going further in speculation, we might estimate the 
magnitude of $m$ as about $s + \sqrt c$ with $s = \max(s_\inp, 
s_\out)$, leading to a complexity of $O(c^{3/2} + sc)$. For 
quasi-optimal algorithms, the term $sc \simeq c^2$ dominates. However, 
as soon as the complexity is at least quadratic in $s$, the dominant 
term is $c^{3/2}$ and the impact on the complexity is then moderated.

@@

\begin{itemize}
\item In which case should we use this package? Mainly for exploration
\item example of $p$-adic differential equations (especially for $p=2$)
\end{itemize}

\subsection{Exploration}

@@

We can then see that up to $63,$ the behaviour
of the loss in relative precision is the same
as for the coefficient in $x^{31}$: \textbf{no loss
occured}. This happens for \ZpCR as for \ZpLC,
and thus, \ZpCR seems to attain the optimal precision.
consequently, those simple manipulations
hint at a different behaviour for the loss 
in precision in at least some special cases of
of isogenies defined over $\mathbb{Q}$
but not necessarily over $\mathbb{Z}_p.$

>>>>

If provably correct results are desired, interval arithmetic provides an alternative to floating point.  Instead of just tracking an approximation to the answer, the package also tracks a radius within which the true result lies.  This method is commonly used for $p$-adic computations since the ultrametric property of $p$-adic fields frequently keeps the radius small.  Computations remain fairly efficient with this approach, but numerical instability can still lead to dramatic losses in precision (see Section \ref{sec:demo} for many examples).

@@

Unfortunately, the latter condition cannot be checked on the matrix 
$M_t$ (because of reductions). 
However it is possible (and easy) to check whether the weaker condition 
$d\Phi_{t,\ttv}(K_t) \subsetneq p^N \Zp$. This checking is achieved by 
the method
{\color{method}\verb?is_precision_capped?} (provided by our package) which returns true 
if $d\Phi_{t,\ttv}(K_t) = p^N \Zp$. As a consequence, when this method
answers \textsc{false}, the absolute precision computed by the software 
is likely optimal.
\end{rmk}

@@

In the \ZpLC framework, creating a variable has a cost $O(m)$. The case 
of deletion is more difficult to handle. However, by the temporal 
locality principle, it seems safe to assume that it is not the 
bottleneck (which is the case in practice). Therefore, when executed 
with the \ZpLF mecanism, the cost of our algorithm is expected to be 
roughly $O(mc)$. Going further in speculation, we might estimate the 
magnitude of $m$ as about $s + \sqrt c$ with $s = \max(s_\inp, 
s_\out)$, leading to a complexity of $O(c^{3/2} + sc)$. For 
quasi-optimal algorithms, the term $sc \simeq c^2$ dominates. However, 
as soon as the complexity is at least quadratic in $s$, the dominant 
term is $c^{3/2}$ and the impact on the complexity is then limited.

@@

The package \ZpL provides powerful tools (based on Automatic 
Differentiation) to track precision in the $p$-adic setting. In many 
concrete situations, it heavily outdoes usual interval arithmetics 
which is currently widely used by Computer Algebra softwares (see
\S \ref{sec:demo} for many examples).
The impact of complexity is controlled but nevertheless non-negligible 
(see \S \ref{ssec:complexity}). For this reason, it is not really 
reasonable to believe that a fast algorithm could rely \emph{directly} 
on the machinery proposed by \ZpL. At least for now, bringing together 
rapidity and stability still requires a substantial human contribution 
and a careful special study of all parameters.

Nevertheless, we do believe that \ZpL can be extremely helpful to 
whoever wants to design a fast and stable $p$-adic algorihm for a couple 
of reasons. First, it provides solutions 
to automatically detect which steps of a given algorithm are stable and 
which ones are not. This way, it highlights the parts of the algorithm 
on which the researcher has to concentrate its efforts.
Second, recall that a classical strategy to improve stability consists 
in working internally at higher precision. The point is that adjusting 
at best the internal precision (\emph{i.e.} finding the good compromise 
between efficiency and accuracy) is not an easy task in general. It 
turns out that the knowledge of the number of diffused digits of 
precision gives very useful hints in this direction. Indeed, as an 
extreme example, consider the case when there is no diffused digit of 
precision. Then, the optimal precision completely splits over the
variables and there is no need to increase internally the precision. On 
the contrary, when there are many diffused digits of precision, a large 
increment is often required.
Since \ZpL gives a direct access to the number of diffused digits of 
precision, it could be very useful to the designer who is concerned
with the balance between efficiency and accuracy.

\begin{comment}
\subsection{Exploration}

@@

We can then see that up to $63,$ the behaviour
of the loss in relative precision is the same
as for the coefficient in $x^{31}$: \textbf{no loss
occured}. This happens for \ZpCR as for \ZpLC,
and thus, \ZpCR seems to attain the optimal precision.
consequently, those simple manipulations
hint at a different behaviour for the loss 
in precision in at least some special cases of
of isogenies defined over $\mathbb{Q}$
but not necessarily over $\mathbb{Z}_p.$ 
\end{comment}
